Running an open experiment: transparency and reproducibility in soil and ecosystem

### 2016-02-04 manuscript_methods/Open_experiment.Rmd
Beefed up intro, tweaked figure

---
title: "Running an open experiment: transparency and reproducibility in soil and ecosystem science"
author: "Ben Bond-Lamberty, Peyton Smith, and Vanessa Bailey"
date: "February 2, 2016"
output: html_document
---

@@

PLOS ONE, ESA, etc. policies

* NSF url
* DOE url
* NASA url
* NSERC, Wellcome trust, other?

@@

Scripted systems for data analysis! E.g. R {R Development Core Team, 2016 #4314}, Kepler (https://kepler-project.org) or Taverna (http://www.taverna.org.uk) record all infrmation about the data processing, analytical process and decisions, and statistical analysis.

@@

@@

>>>>

---
title: 'Running an open experiment: transparency and reproducibility in soil and ecosystem
  science'
author: "Ben Bond-Lamberty, Peyton Smith, and Vanessa Bailey"
date: "February 2, 2016"
output: html_document
---

#### Abstract

Researchers in soil and ecosystem science, and almost every other field, as increasingly being pushed--by funders, journals, governments, and themselves--to ensure transparency and reproducibility of their work. A key part of this effort is a move towards *open data* as a way to fight post-publication data loss, improve data and code quality, and enable powerful meta- and cross-disciplinary analyses. We describe a recent "open experiment", in which we documented every aspect of a soil incubation experiment online, making publicly available raw data, scripts, diagnostics, final analyses, and manuscripts in real time. Our use of tools such as version control, issue tracking, and open-source statistical software ensured data integrity and transparency. There are many avenues to improve scientific reproducibility and data availability, of which is this only one example; we acknowledge that this is not an approach suited for every experiment or situation. Nonetheless, we encourage the communities in our respective fields to consider its advantages, and lead rather than follow with respect to scientific reproducibility, transparency, and data availability.

@@

Journals are adopting increasingly stringent data access and deposition policies, e.g. Scientific Data (http://www.nature.com/sdata/data-policies), PLOS ONE (http://journals.plos.org/plosone/s/data-availability),  {Scientific Data, 2016 #4339}{PLOS ONE, 2014 #4338}, and Science (http://www.sciencemag.org/authors/science-editorial-policies#data-deposition). These policies generally share common assumptions and goals: maximizing access to scientific data; encouraging deposition into structured repositories as opposed to journal supplementary information; and specifying that it is not acceptable for authors to be solely responsible for ensuring data access.

Funders are moving in this direction as well, with organizations such as the U.S. National Science Foundation (http://www.nsf.gov/bfa/dias/policy/dmp.jsp) and the U.K. Wellcome Trust (http://www.wellcome.ac.uk/About-us/Policy/Spotlight-issues/Data-sharing/Data-management-and-sharing/index.htm) requiring explicit data management plans, unfettered reasonable access to primary data, and use of established repositories.

Finally, growing numbers of scientists are pushing for open data on moral/political (why should the public fund scientists if the latter aren't producing demonstrably replicable and reusable results?) grounds, also on purely scientific ones: ensuring reproducibility {Stodden, 2011 #4340}. This latter issue--reproducibility--is at the heart of science {Kuhn, 1962 #4341} and has a number of facets worth considering briefly.

@@

Scripted systems for data analysis! E.g. R {R Development Core Team, 2016 #4314}, Kepler (https://kepler-project.org) or Taverna (http://www.taverna.org.uk) record all infrmation about the data processing, analytical process and decisions, and statistical analysis. This and other best practices of scientific computing {Wilson, 2014 #4337} help produce reproducible, robust, and transparent science.

@@

A subtler data loss issue is the "file drawer problem" {Rosenthal, 1979 #2291}.

@@

TODO - maybe a second figure showing one of the diagnostic plots, partway through the incubation?
